Biologically-Constrained Region Merging for Connectome Reconstruction

### 2018-03-14 papers/eccv2018/0_introduction.tex
pass on related work, use region merging instead of agglomeration

The field of connectomics considers the problem of reconstructing the wiring diagram of the brain at nanometer resolutions to enable new insights into the workings of the brain~\cite{haehn2017scalable,kasthuri2015saturated}. 
%Recent advancements in image acquisition using multi-beam serial-section electron microscopy (sSEM) allow researchers to produce terabytes of image data per hour~\cite{hildebrand2017whole}. 
%It is infeasible for domain experts to manually reconstruct this vast amount of image data~\cite{haehn2014design}. 
State-of-the-art automatic 3D reconstruction approaches use pixel-based convolutional neural networks (CNNs) and watershed transforms to generate initial segmentation, followed by agglomeration steps~\cite{seymour2016rhoananet,lee2015recursive,nunez2014graph,parag2017anisotropic,ronneberger2015u,zlateski2015image}.
For the agglomeration, most methods do not use prior biological knowledge of the 3D shape of neural structures.
Reported on small-sized volumes, a simple threshold-based mean agglomeration method can achieve superhuman result~\cite{lee2017superhuman}.
However, it is unclear if that is the case for larger volumes.

%\TODO{the selling point is to add biological constraint}
%We present a \textit{top-down graph-based} method that builds on the outputs of bottom-up pixel-based segmentation approaches. 
We present a new agglomeration framework that imposes both local and global biological constraints onto the output segmentation to match the underlying structure of neuronal processes.
We first extract skeletons from labels in the input segmentation and generate a simplified graph (Fig.~\ref{fig:teaser_pipeline}a). 
Next, we train a CNN classifier to learn biological constraints on the local structural shape of a neuron (Fig.~\ref{fig:teaser_pipeline}b). 
This classifier produces probabilities that two segments belong to the same neuron based on the input shapes of the segments.
These probabilities correspond to edge weights in our graph extracted from the input segmentation.
We then use a graph optimization algorithm to partition the graph into a final improved reconstruction by enforcing domain-specific topological constraints from biology (Fig.~\ref{fig:teaser_pipeline}c).

>>>>

The field of connectomics considers the problem of reconstructing the wiring diagram of the brain at nanometer resolutions to enable new insights into the workings of the brain~\cite{haehn2017scalable,kasthuri2015saturated}. 
%Recent advancements in image acquisition using multi-beam serial-section electron microscopy (sSEM) allow researchers to produce terabytes of image data per hour~\cite{hildebrand2017whole}. 
%It is infeasible for domain experts to manually reconstruct this vast amount of image data~\cite{haehn2014design}. 
State-of-the-art automatic 3D reconstruction approaches use pixel-based convolutional neural networks (CNNs) and watershed transforms to generate initial over-segmentation, followed by region merging steps~\cite{seymour2016rhoananet,lee2015recursive,nunez2014graph,parag2017anisotropic,ronneberger2015u,zlateski2015image}.
For the region merging, most methods do not use prior biological knowledge of the 3D shape of neural structures.
Reported on small-sized volumes, a simple threshold-based mean agglomeration method can achieve superhuman result~\cite{lee2017superhuman}.
However, it is unclear if that is the case for larger volumes.
\TODO{sth. about figure 1.}

%\TODO{the selling point is to add biological constraint}
%We present a \textit{top-down graph-based} method that builds on the outputs of bottom-up pixel-based segmentation approaches. 
We present a new region merging framework that imposes both local and global biological constraints onto the output segmentation to match the underlying structure of neuronal processes.
We first extract skeletons from labels in the input segmentation and generate a simplified graph (Fig.~\ref{fig:teaser_pipeline}a). 
Next, we train a CNN classifier to learn biological constraints on the local structural shape of a neuron (Fig.~\ref{fig:teaser_pipeline}b).
This classifier produces probabilities that two segments belong to the same neuron based on the input shapes of the segments.
These probabilities correspond to edge weights in our graph extracted from the input segmentation.
We then use a graph optimization algorithm to partition the graph into a final improved reconstruction by enforcing domain-specific topological constraints from biology (Fig.~\ref{fig:teaser_pipeline}c).

### 2018-03-14 papers/eccv2018/1_related_work.tex
pass on related work, use region merging instead of agglomeration

\subsubsection{Pixel-Based Segmentation}
A significant amount of connectomics research considers the problem of extracting segmentation information at the voxel level from raw EM images.

First, intermediate representations like boundary or affinity maps are generated. Some methods use a random forest populated with hand-designed features based on image characteristics around a given pixel to predict membrane probabilities~\cite{kaynig2015large}.
Most recent advancements use convolutional neural networks since machine-learned features outperform hand-designed ones~\cite{bogovic2013learned}.
Originally, networks predicted membrane probabilities per each image slice using 2D convolutions~\cite{seymour2016rhoananet,ronneberger2015u,ciresan2012deep,jain2010boundary,kaynig2015large,amelio_segmentation}.
However, recent research extends these convolutions to include the z-dimension improving performance~\cite{lee2015recursive,parag2017anisotropic,cciccek20163d,turaga2010convolutional}.
Many of these networks produce probabilities for the affinity between two voxels (i.e., the probability that adjacent voxels belong to the same neuron). 
The MALIS cost function is specifically designed for generating affinities that produce good segmentations~\cite{briggman2009maximin}. 
More recently, flood-filling networks produce segmentations by training an end-to-end neural network that goes from EM images directly to label volumes~\cite{januszewski2016flood}. 
These networks produce impressive accuracies but at a high computational cost.
Data augmentation in both training and testing can increase the prediction accuracy in boundary detection~\cite{zeng2017deepem3d}.
Lee et al. surpass the estimated human accuracy on the SNEMI3D dataset with a variant of the 3D U-Net architecture~\cite{lee2017superhuman}.

Then, these intermediate representations are transformed into segmentation.
Some early techniques apply computationally expensive graph partitioning algorithms with a single node per superpixel~\cite{andres2012globally}.
Several pixel-based approaches generate probabilities that neighboring pixels belong to the same neuron.
Often a watershed algorithm will then cluster pixels into super-pixels~\cite{zlateski2015image}.

\subsubsection{Agglomeration}
\TODO{rule-based}
Some methods rely solely on the accuracy of the predicted affinities and agglomerate hierarchically based on the mean affinity between supervoxels~\cite{lee2017superhuman,funke2017deep}.

\TODO{classification-based}
Given the initial segmentation, some methods train random-forest classifiers to produce the final segmentations~\cite{seymour2016rhoananet,nunez2014graph,parag2017anisotropic,zlateski2015image,10.1371/journal.pone.0125825}
extract different features
~\cite{jain2011learning,bogovic2013learned}.

\TODO{optimization-based}
Pape et al. present a scalable multicut algorithm to partition superpixels~\cite{beier2017multicut}.

In our work, we combine classification and optimization.
%\subsubsection{Error-correction methods}
%Additional research builds on top of these region-based methods to correct errors in the segmentation either using human proofreading~\cite{haehn2014design,haehn2017guided,mojo2} or automatic methods~\cite{rolnick2017morphological,error_correction_using_CNN}.
%However, to our knowledge, our method is the first to extract a 3D graph from an input segmentation for a true top-down\TODO{?} error correction approach. 
%This allows us to enforce domain-specific biology constraints and use efficient graph partitioning algorithms. 
% Many segmentation and clustering algorithms use graph partitioning techniques~\cite{andres2012globally} or normalized cuts for traditional image segmentation~\cite{kappes2016higher,shi2000normalized,tatiraju2008image}.
% Even though graph partitioning is an NP-Hard problem~\cite{demaine2006correlation} there are several useful multicut heuristics that provide good approximations with reasonable computational costs~\cite{horvnakova2017analysis}. 
%We use the method of Keuper et al.~\cite{keuper2015efficient} to partition the extracted graph into the final neural reconstruction.

>>>>

%\noindent\textbf{Pixel-Based Segmentation}
\subsubsection{Pixel-Based Segmentation}
A significant amount of connectomics research considers the problem of extracting segmentation information at the voxel level of EM images.
First, intermediate representations like boundary, affinity or binary segmentation maps are generated from voxels.
For the boundary probabilities, random forest with hand-designed features~\cite{kaynig2015large}, 2D and 3D convolution networks ~\cite{bogovic2013learned,seymour2016rhoananet,ronneberger2015u,ciresan2012deep,jain2010boundary,amelio_segmentation}.
Later, the affinity between each voxel and its six neighors are widely used~\cite{lee2015recursive,parag2017anisotropic,cciccek20163d,turaga2010convolutional,lee2017superhuman}. 3D U-Net architecture becomes popular~\cite{cciccek20163d} and MALIS cost function is designed to re-weight affinity prediction by their contribution to the segmentation error~\cite{briggman2009maximin}.
More recently, flood-filling networks~\cite{januszewski2016flood} produce binary segmentations from raw pixels with a recurrent convolution network at a high computational cost.
Orthogonal to the representation, model averaging~\cite{zeng2017deepem3d} and data augmentation~\cite{lee2017superhuman} methods can further improve the performance, where Lee et al.~\cite{lee2017superhuman} surpass the estimated human accuracy on the SNEMI3D dataset.

Then, clustering techniques transform these intermediate representations into segmentations.
Some early techniques apply computationally expensive graph partitioning algorithms with a single node per superpixel~\cite{andres2012globally}.
Several pixel-based approaches generate probabilities that neighboring pixels belong to the same neuron.
Often a watershed algorithm will then cluster these pixels into super-pixels~\cite{zlateski2015image}.

%\noindent\textbf{Region Merging}
\subsubsection{Region Merging}
Region merging methods can be categorized by the similairity metric between adjacent segments and the merging strategy.
For the similarity metric, Lee et al. and Funke et al. rely solely on the accuracy of the predicted affinities and define the metric as the mean affinity between segments~\cite{lee2017superhuman,funke2017deep}.
Classification-based method predicts the probability to merge two segments from handcrafted~\cite{seymour2016rhoananet,nunez2014graph,parag2017anisotropic,zlateski2015image,10.1371/journal.pone.0125825,jain2011learning} and learned features~\cite{bogovic2013learned}.

For the merging stategy, most methods use variants of hierarchical agglomeration methods~\cite{seymour2016rhoananet,nunez2014graph,parag2017anisotropic,zlateski2015image,10.1371/journal.pone.0125825,bogovic2013learned} to greedily merge a pair of regions at a time.
Jain et al. formulates the agglomeration as a reinforecement learning problem ~\cite{jain2011learning} and Pape et al. present a scalable multicut algorithm to partition superpixels with global optimization~\cite{beier2017multicut}.

% \subsubsection{Error-correction methods}
% Additional research builds on top of these region-based methods to correct errors in the segmentation either using human proofreading~\cite{haehn2014design,haehn2017guided,mojo2} or automatic methods~\cite{rolnick2017morphological,error_correction_using_CNN}.
% Many segmentation and clustering algorithms use graph partitioning techniques~\cite{andres2012globally} or normalized cuts for traditional image segmentation~\cite{kappes2016higher,shi2000normalized,tatiraju2008image}.
% Even though graph partitioning is an NP-Hard problem~\cite{demaine2006correlation} there are several useful multicut heuristics that provide good approximations with reasonable computational costs~\cite{horvnakova2017analysis}. 
%We extract a 3D graph from an input segmentation for a true top-down error correction approach. 
%This allows us to enforce domain-specific biological constraints while using graph partitioning algorithms. 
%We use the method of Keuper et al.~\cite{keuper2015efficient} to partition the extracted graph into the final neural reconstruction.

%\subsubsection{Error-correction methods}
%Additional research builds on top of these region-based methods to correct errors in the segmentation either using human proofreading~\cite{haehn2014design,haehn2017guided,mojo2} or automatic methods~\cite{rolnick2017morphological,error_correction_using_CNN}.
%However, to our knowledge, our method is the first to extract a 3D graph from an input segmentation for a true top-down\TODO{?} error correction approach. 
%This allows us to enforce domain-specific biology constraints and use efficient graph partitioning algorithms. 
% Many segmentation and clustering algorithms use graph partitioning techniques~\cite{andres2012globally} or normalized cuts for traditional image segmentation~\cite{kappes2016higher,shi2000normalized,tatiraju2008image}.
% Even though graph partitioning is an NP-Hard problem~\cite{demaine2006correlation} there are several useful multicut heuristics that provide good approximations with reasonable computational costs~\cite{horvnakova2017analysis}. 
%We use the method of Keuper et al.~\cite{keuper2015efficient} to partition the extracted graph into the final neural reconstruction.

### 2018-03-14 papers/eccv2018/2_method.tex
pass on related work, use region merging instead of agglomeration

% !TEX root =  paper.tex
\section{Method}

\begin{figure}[t!]
	\centering
	\includegraphics[width=\linewidth]{./figures/teaser-image.png}
	\caption{Our proposed framework uses both local and global biological constraints to improve an input segmentation. We generate a simplified graph from an input segmentation (a), train a classifier to learn neuron shape constraints locally (b), and perform graph partitioning to enforce domain-specific properties globally.\TODO{Is it possible to have the weights next to the lines?}}
	\label{fig:teaser_pipeline}
\end{figure}
%\begin{figure*}[t!]
%	\begin{center}
%		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/multicut_input.png}}
%		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/pre-multicut.png}}
% 		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/multicut-graph.png}}
% 		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/post-multicut.png}}
% 	\end{center}
% 	\caption{Outline of our approach, from left to right: result of the pixel-based segmentation and agglomeration algorithm; segments of several neurons from the initial segmentation; extracted skeletonized network of those segments; improved 3D reconstruction of the selected segments after graph construction and partitioning with constraints.}
% 	\label{fig:overview}
% \end{figure*}

There are two types of errors that can occur in connectomics segmentations: \textit{split errors} and \textit{merge errors}. 
In a split error, two segments should have been merged into one neuronal process (Fig.~\ref{fig:improved-reconstruction}). 
In a merge error, one segment corresponds to more than one neuron.
Generally, it is much more difficult to correct merge errors than to correct split errors, as the space of possible split proposals grows quickly~\cite{parag2015properties}.
Thus, most reconstruction approaches are tuned towards over-segmentation with many more split than merge errors. 
The input of our method is the over-segmentation of EM image volumes from state-of-the-art pipelines.

>>>>

% !TEX root =  paper.tex
\begin{figure}[t!]
	\centering
	\includegraphics[width=\linewidth]{./figures/teaser-image.png}
	\caption{Our proposed framework uses both local and global biological constraints to improve an input segmentation. We generate a simplified graph from an input segmentation (a), train a classifier to learn neuron shape constraints locally (b), and perform graph partitioning to enforce domain-specific properties globally.\TODO{Is it possible to have the weights next to the lines?}}
	\label{fig:teaser_pipeline}
\end{figure}
\section{Method}
%\begin{figure*}[t!]
%	\begin{center}
%		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/multicut_input.png}}
%		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/pre-multicut.png}}
% 		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/multicut-graph.png}}
% 		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/post-multicut.png}}
% 	\end{center}
% 	\caption{Outline of our approach, from left to right: result of the pixel-based segmentation and agglomeration algorithm; segments of several neurons from the initial segmentation; extracted skeletonized network of those segments; improved 3D reconstruction of the selected segments after graph construction and partitioning with constraints.}
% 	\label{fig:overview}
% \end{figure*}

%There are two types of errors that can occur in connectomics segmentations: \textit{split errors} and \textit{merge errors}. 
%In a split error, two segments should have been merged into one neuronal process (Fig.~\ref{fig:improved-reconstruction}). 
%In a merge error, one segment corresponds to more than one neuron.
%Generally, it is much more difficult to correct merge errors than to correct split errors, as the space of possible split proposals grows quickly~\cite{parag2015properties}.
%Thus, most reconstruction approaches are tuned towards over-segmentation with many more split than merge errors. 
%The input of our method is the over-segmentation of EM image volumes from state-of-the-art pipelines.

### 2018-03-14 papers/eccv2018/paper.tex
First pass of fixes.
pass on related work, use region merging instead of agglomeration

\authorrunning{ECCV-18 submission ID \ECCV18SubNumber}
%\title{Scalable Graph-Based Neural \\ Reconstruction for Connectomics}
\title{Biologically-Constrained 3D Agglomeration for Connectome Reconstruction}

@@

\begin{abstract}
% maybe better to cut to the point
%Advancements in electron microscopy image acquisition have created connectomics datasets terabytes in size making manual reconstruction of neuronal structures infeasible.
%In the past decade in Connectomics, much progress has been made to classify neural membranes or affinities from the raw pixel and to watershed transform them into initial segments.
%Current state-of-the-art automatic methods classify neural membranes or affinities from the raw pixel followed by agglomeration methods to create full neuron reconstructions.
%An agglomerative segmentation algorithm merges initial pair of clusters.
Current state-of-the-art connectome reconstruction pipelines have two major steps: initial pixel-based segmentation with affinity prediction and watershed transform, and refined segmentation with agglomeration.
%In the past deTraditionally, neuron segmentation from 3D electron microscopy (EM)
For the agglomeration, most methods neglect geometric and topological constraints that are inherent in the graph structure of neural wiring diagrams.

In this work, we use a biologically-constrained graph-based method to better reconstruct neural pathways.
Given the initial segmentation, we first create a graph with nodes corresponding to segment labels and edges identifying potential locations of split errors based on the skeleton structure of the segmentation.
Next, we train a convolutional neural network to capture the geometric properties of segment connections and produce edge weights as the probability of merging two nodes.
Finally, we formulate the agglomeration as a lifted multicut partitioning problem with topological constraints to avoid cycles.
We demonstrate the performance of our approach on multiple real-world connectomics datasets with an average split variation of information improvement of \TODO{XX}\%.
\keywords{connectomics, biological constraints, agglomeration, skeletonization, deep learning, lifted multicut}
\end{abstract}

>>>>

\authorrunning{ECCV-18 submission ID \ECCV18SubNumber}
%\title{Scalable Graph-Based Neural \\ Reconstruction for Connectomics}
%\title{Biologically-Constrained 3D Agglomeration for Connectome Reconstruction}
\title{Biologically-Constrained Region Merging for Connectome Reconstruction}

@@

\begin{abstract}
% maybe better to cut to the point
%Advancements in electron microscopy image acquisition have created connectomics datasets terabytes in size making manual reconstruction of neuronal structures infeasible.
%In the past decade in Connectomics, much progress has been made to classify neural membranes or affinities from the raw pixel and to watershed transform them into initial segments.
%Current state-of-the-art automatic methods classify neural membranes or affinities from the raw pixel followed by agglomeration methods to create full neuron reconstructions.
%An agglomerative segmentation algorithm merges initial pair of clusters.
Current state-of-the-art connectome reconstruction pipelines have two major steps: initial pixel-based segmentation with affinity prediction and watershed transform, and refined segmentation by merging oversegmented regions.
%In the past deTraditionally, neuron segmentation from 3D electron microscopy (EM)
For the region merging, most methods neglect geometric and topological constraints that are inherent in the graph structure of neural wiring diagrams.
In this work, we use a biologically-constrained graph-based method to better reconstruct neural pathways.
Given the initial segmentation, we first create a graph with nodes corresponding to segment labels and edges identifying potential locations of split errors based on the skeleton structure of the segmentation.
Next, we train a convolutional neural network to capture the geometric properties of segment connections and produce edge weights as the probability of merging two nodes.
Finally, we formulate the region merging as a lifted multicut partitioning problem with topological constraints to avoid cycles.
We demonstrate the performance of our approach on multiple real-world connectomics datasets with an average split variation of information improvement of \TODO{XX}\%.
\keywords{connectomics, biological constraints, region merge, skeletonization, deep learning, lifted multicut}
\end{abstract}
